---
title: "Assignment 2"
author: "Marine Courtin"
date: "9 octobre 2017"
output: html_document
---


## Exercise 1. Writing R functions


#### Question a.

+ __Loading functions__

    This chunk allows us to load functions that we've written in a separate R file :

```{r}
source("functions.R")
```

+ __Test on the existence of a column__

    It is rather easy to test whether a column exists. First we attribute the value returned by the column of the data frame. Then, we test `is.null(value)`. The boolean returned is True when the column does not exist and False when it does. It's also possible to test for the absence of the column using the __!__ operator.

+ __Test on the type of value of a vector__

    We can test the type of the data contained in our vector. Using `is.numeric(vector)` will return True if all values in our vector are numeric. As a reminder, elements of R vectors must be of the same type. If we do not respect this condition when creating the vector - for example declaring `vec <- c('mot`, 5)` - R automatically chooses a datatype in our place (here 5 would be converted to string).


Here's what we get when we apply the function : 

```{r}
print(sum_column(iris, "Sepal.Length"))
print(sum_column(iris, "Species"))
print(sum_column(warpbreaks, "breaks"))
```


#### Question b. 

+ __Boucler sur les valeurs d'un vecteur__

In order to add up the values of our vector, we loop on the elements and iteratively add their value to a variable. We first need to find out the number of elements using `length(vector)`. Then we create a loop which respects the `for (condition) {do}` construction. Our condition specifies the possibles values of the element's indexes : from 1 up to n, if we have n elements in our vector. We simply add the i*th* element of our vector to the total sum each time.


```{r}
print(my_sum(iris$Sepal.Length))
print(my_sum(iris$Species))
print(my_sum(warpbreaks$breaks))
```

#### Question c.

+ __Testing a combination of conditions__

This time, we want to use our `my_sum()` function inside another function `sum_divided_by()`. We give it two arguments : a vector of number x, and a number k by which we'll divide x. We initialize a NULL value to return in case our arguments do not respect these conditions. 
    
Here's the result :

```{r}
print(sum_divided_by(iris$Sepal.Length, 12))
print(sum_divided_by(iris$Species, 22))
print(sum_divided_by(iris$Sepal.Length, "Not numeric"))
print(sum_divided_by(warpbreaks$breaks, -12))
```

#### Question d.

+ __Calculating the mean of a vector__

We want to write a function which will calculate the mean of a vector. In order to do that, we can use the `sum_divided_by()` function we've juste created, and simply provide the length of our vector as the k argument.

We do not need to initialize the result value as NULL in this new function, as it is already part of the output produced by `sum_divided_by()`. If x is not a vector of number, `sum_divided_by()` will return NULL.

```{r}
print(my_mean(iris$Sepal.Length))
print(my_mean(iris$Species))
print(my_mean(warpbreaks$breaks))
```

